---
title: "Repro Research - Project 1"
output: md_document
---

Reading in the data, loading libraries

```{r}
activity<-read.csv("activity.csv")
library(ggplot2)
```

#Q1.What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

Calculate the total number of steps taken per day
If you do not understand the difference between a histogram and a barplot, research the difference between them. 
Make a histogram of the total number of steps taken each day
Calculate and report the mean and median of the total number of steps taken per day

Finding steps per day:

```{r}
View(activity)
perdaysteps<-tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

Plotting histogram of steps/day

```{r}
hist(perdaysteps, xlab="Steps per day", main="Histogram of Steps per day")
```

Finding mean and median

```{r}
meansteps<-mean(perdaysteps)
mediansteps<-median(perdaysteps)
print("Mean and median respectively are "); meansteps; mediansteps
```

#Q2. What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
# preprocessing data for plot
steps_by_interval <- aggregate(steps ~ interval, activity, mean)

# create a time series plot 
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

```{r}
# find row with max of steps
max_steps_row <- which.max(steps_by_interval$steps)

# find interval with this max
steps_by_interval[max_steps_row, ]
```


#Q.3. Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? 
What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
sum(is.na(activity))
```

There are 2304 missing values

Filling in the NA's with mean for 5 minute interval

```{r}
data_imputed <- activity
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
# calculate  total number of steps taken each day
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)

#Plotting Histogram
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```

```{r}
# get mean and median of imputed data
mean(df_imputed_steps_by_day$steps)

median(df_imputed_steps_by_day$steps)
```

There is a slight difference in Median


#Q.4 Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

Function to determine if weekday
```{r}
week_day <- function(date_val) {
    wd <- weekdays(as.Date(date_val, '%Y-%m-%d'))
    if  (!(wd == 'Saturday' || wd == 'Sunday')) {
        x <- 'Weekday'
    } else {
        x <- 'Weekend'
    }
    x
}
```

Plotting the graph
```{r}
# Apply the week_day function and add a new column to activity dataset
activity$day_type <- as.factor(sapply(activity$date, week_day))

#load the ggplot library
library(ggplot2)

# Create the aggregated data frame by intervals and day_type
steps_per_day_impute <- aggregate(steps ~ interval+day_type, activity, mean)

# Create the plot
plt <- ggplot(steps_per_day_impute, aes(interval, steps)) +
    geom_line(stat = "identity", aes(colour = day_type)) +
    theme_gray() +
    facet_grid(day_type ~ ., scales="fixed", space="fixed") +
    labs(x="Interval", y=expression("No of Steps")) +
    ggtitle("No of steps Per Interval by day type")
print(plt)
```


